\section{Introduction}

Analytical pipelines in scientific, engineering, policy settings routinely produce discrete outcomes such as a selected route, a diagnostic label, a risk classification, a policy recommendation. These outcomes depend not only on the input data and the computational procedure, but also on representational choices, including how the data is encoded, what features are weighted, which aggregation rules are applied. Small changes to such choices can induce qualitatively different outcomes, even when the underlying data and computation are unchanged.

These dependencies are real and consequential, yet they are rarely recorded as first-class objects of study. In standard practice, a pipeline is run under one representation, a result is reported, the sensitivity of that result to representational alternatives goes unexamined. When instabilities surface, they are typically discovered after deployment or during post-hoc audits, because no infrastructure existed to make them visible earlier.

This paper introduces a diagnostic framework for making representational dependence observable. The central object is a \emph{decision-valued map}, a mapping from a family of representations to discrete decision identities, evaluated under a fixed data snapshot and a fixed computational engine. By materializing this map across controlled representational variation, it becomes possible to observe directly which outcomes persist under representation changes and where boundaries form.

DecisionDB is the system that implements this protocol. It logs snapshots, representations, engine runs, decision identities using content-addressed identifiers and immutable artifacts. It supports representational sweeps through systematic variation of declared representation parameters, replay verification through deterministic recovery of decision identifiers from persisted artifacts, post-hoc audit of the full provenance chain.

The approach is demonstrated on a graph routing problem. Fixing a graph snapshot and a shortest-path engine, two representation parameters that control edge-cost construction are swept. One parameter preserves route identity across its tested range; the other induces a discrete change in route identity at a specific threshold. Replay verification confirms that persisted identifiers are deterministically recoverable.

The contribution is not a model, a learning rule, or an optimization procedure. It is an infrastructure for observing what happens to discrete outcomes when representations change.
